---
title: "Maladaptive plastic responses of flowering time to geothermal heating (Cerastium 2)"
subtitle: "Logger data preparation"
author : "Alicia Valdés"
output:
  pdf_document:
    toc: yes
    toc_depth: 4
  html_notebook:
    toc: yes
    toc_depth: '4'
    latex_engine: xelatex
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r load packages, include=FALSE}
library(tidyverse)
library(ggthemes)
library(knitr)
library(gridExtra)
library(DHARMa)
library(RColorBrewer)
library(broom)
library(ggpubr)
library(jtools)
library(kableExtra)
library(ggeffects)
library(MuMIn)
library(MASS)
library(segmented)
library(effects)
library(purrr)
library(lubridate)
library(ggforce)
library(car)
library(ggrepel)
```

```{r Define ggplot themes and palettes, include=FALSE}
my_theme <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(legend.position="none")+theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
my_theme_legend <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
myPalette <- colorRampPalette(brewer.pal(11, "YlOrRd"))
```

```{r, include=FALSE}
set_summ_defaults(digits = 3)
```

# Read list of all plants with temperature measured in 2018 (including Cerastium and other species)

```{r}
ids_temp_allsps<-read.table("data/edited/ids_temp_allsps.csv",
                            header=T,sep=",",dec=".") 
ids_temp_allsps$temp_term<-ids_temp_allsps$temp
ids_temp_allsps$temp<-NULL
head(ids_temp_allsps)
```

temp\_term = soil temperature measured with a thermometer at the time of marking in 2018, measured next to each plant at a depth of 10 cm.

# Read list of plants with loggers (including Cerastium and other species)

```{r}
loggers<-read.table("data/edited/ibuttons_AV_allsps.csv",
  header=T,sep=",",dec=".") 
head(loggers)
```

# Merge both

Keeping records in both lists (i.e. plants with temperature measured with thermometer and with a logger).

```{r message=FALSE, warning=FALSE}
temp_loggers<-ids_temp_allsps%>%
  inner_join(loggers)
head(temp_loggers)
```

```{r fig.height=3, fig.width=4}
ggplot(temp_loggers,aes(x=factor(above_below)))+geom_bar()
```

79 above and 143 below.

# Import raw iButton data

```{r}
all_logger_data <- list()
for(filename in list.files("data/edited/RAW_iButton_data_to_use",
  full.names=T)) {
   all_logger_data[[length(all_logger_data) + 1]] <- read.table(filename,
                                                header=T,fill=T,sep=",",
                                                dec=",",skip=14) 
   all_logger_data[[length(all_logger_data)]]$logger_nr <-substr(filename,
                                                                 start=37,
                                                                 stop=100)
}
logger_data <- do.call(rbind, all_logger_data)%>%
  rownames_to_column("datetime")%>%
  rename(unit=Date.Time,value_int=Unit,value_dec=Value)
logger_data<-logger_data[c(1,3:5)]%>%
  mutate(temp=as.numeric(ifelse(is.na(value_dec),
                                value_int,
                                paste(value_int,value_dec,sep="."))))
logger_data<-logger_data[c(1,4:5)]%>%
  mutate(datetime=as.POSIXct(logger_data$datetime,
                             format="%d.%m.%Y %H:%M:%S",tz=Sys.timezone()))
logger_data<-logger_data%>%
  mutate(datetime=datetime %m+% years(1))%>%
  # Add one year to dates, as they were wrong - CHECK SOME CASES WERE NOT!
  mutate(logger_nr=as.numeric(gsub(".csv","",logger_nr)))

# Read individually some files that gave problems
log339<-read.table("data/edited/RAW_iButton_data_to_use_problematic/339.csv",
           header=T,fill=T,sep=",",dec=",",skip=14)
log344<-read.table("data/edited/RAW_iButton_data_to_use_problematic/344.csv",
           header=T,fill=T,sep=",",dec=",",skip=14)
log352<-read.table("data/edited/RAW_iButton_data_to_use_problematic/352.csv",
           header=T,fill=T,sep=",",dec=",",skip=14)
log383<-read.table("data/edited/RAW_iButton_data_to_use_problematic/383.csv",
           header=T,fill=T,sep=",",dec=",",skip=14)
log431<-read.table("data/edited/RAW_iButton_data_to_use_problematic/431.csv",
           header=T,fill=T,sep=",",dec=",",skip=14)
log585<-read.table("data/edited/RAW_iButton_data_to_use_problematic/585.csv",
           header=T,fill=T,sep=",",dec=",",skip=14)
log641<-read.table("data/edited/RAW_iButton_data_to_use_problematic/641.csv",
           header=T,fill=T,sep=",",dec=",",skip=14)
log654<-read.table("data/edited/RAW_iButton_data_to_use_problematic/654.csv",
           header=T,fill=T,sep=",",dec=",",skip=14)
log663<-read.table("data/edited/RAW_iButton_data_to_use_problematic/663.csv",
           header=T,fill=T,sep=",",dec=",",skip=14)
log339$logger_nr<-339
log344$logger_nr<-344
log352$logger_nr<-352
log383$logger_nr<-383
log431$logger_nr<-431
log585$logger_nr<-585
log641$logger_nr<-641
log654$logger_nr<-654
log663$logger_nr<-663

logger_data_problematic<-rbind(log339,log344,log352,log383,log431,
                               log585,log641,log654,log663)
logger_data_problematic<-logger_data_problematic%>%
  rename(datetime=Date.Time)%>%
   mutate(temp=as.numeric(ifelse(is.na(value_dec),
                                value_int,
                                paste(value_int,value_dec,sep="."))))
logger_data_problematic<-logger_data_problematic[c(1,5:6)]%>%
  mutate(datetime=as.POSIXct(logger_data_problematic$datetime,
                             format="%d.%m.%Y %H:%M:%S",
                             tz=Sys.timezone()))%>%
  mutate(datetime=datetime %m+% years(1))
  # Add one year to dates, as they were wrong - CHECK SOME CASES WERE NOT!

logger_data<-rbind(logger_data,logger_data_problematic)%>%
  inner_join(temp_loggers)
head(logger_data) 
```

```{r}
nrow(subset(logger_data,is.na(datetime)))
```

There are some NAs for datetime (70 rows - it seems that they were not able to be converted to POSIXct) --\> We can live with those

# Read coordinates of loggers and add to data:

```{r}
coords<-read.table("data/edited/coords_AV.csv",
  header=T,sep=",",dec=".") 
logger_data<-logger_data%>%
  left_join(coords)
head(logger_data)
```

# Plot locations of loggers in each plot, calculate distance matrices and make pairs

For each plot, pair each aboveground logger with the closest belowground logger

```{r message=FALSE, warning=FALSE}
logger_data%>%
  filter(plot=="HC1")%>%
  group_by(plot,logger_nr,above_below)%>%
  summarise(X=first(X),Y=first(Y))%>%
  ggplot(aes(x=X,y=Y,color=above_below,label=logger_nr))+geom_point()+
  geom_text_repel()+ggtitle("Plot HC1")+
  scale_y_continuous(breaks = seq(0,13,by=1)) +
  scale_x_continuous(breaks = seq(0,13,by=1)) +
  coord_fixed(ylim=c(0, 9),xlim=c(0, 12))
```

Distance matrix:

```{r message=FALSE, warning=FALSE}
logger_data%>%
  filter(plot=="HC1")%>%
  group_by(logger_nr)%>%
  summarise(X=first(X),Y=first(Y))%>%
  column_to_rownames(var="logger_nr")%>%
  dist()%>%
  as.matrix()%>%
  # as.data.frame()%>%
  # rownames_to_column(var="logger_nr")%>%
  # pivot_longer(cols=-logger_nr,names_to="logger_nr_2",values_to="distance")
  round(digits=2)
```

Pairs: 103-105 and 104-101

```{r}
logger_data<-logger_data%>%
  mutate(pair=ifelse(logger_nr==103|logger_nr==105,1,
                     ifelse(logger_nr==104|logger_nr==101,2,NA)))%>%
  mutate(dist=ifelse(pair==1,0.51,
                     ifelse(pair==2,5.65,NA)))
```

```{r message=FALSE, warning=FALSE}
logger_data%>%
  filter(plot=="HC2a")%>%
  group_by(plot,logger_nr,above_below)%>%
  summarise(X=first(X),Y=first(Y))%>%
  ggplot(aes(x=X,y=Y,color=above_below,label=logger_nr))+geom_point()+
  geom_text_repel()+ggtitle("Plot HC2a")+
  scale_y_continuous(breaks = seq(0,13,by=1)) +
  scale_x_continuous(breaks = seq(0,13,by=1)) +
  coord_fixed(ylim=c(0, 3),xlim=c(0, 9))
```

Distance matrix:

```{r message=FALSE, warning=FALSE}
logger_data%>%
  filter(plot=="HC2a")%>%
  group_by(logger_nr)%>%
  summarise(X=first(X),Y=first(Y))%>%
  column_to_rownames(var="logger_nr")%>%
  dist()%>%
  as.matrix()%>%
  # as.data.frame()%>%
  # rownames_to_column(var="logger_nr")%>%
  # pivot_longer(cols=-logger_nr,names_to="logger_nr_2",values_to="distance")
  round(digits=2)
```

Pairs: 108-113 and 111-115

```{r}
logger_data<-logger_data%>%
  mutate(pair=ifelse(logger_nr==108|logger_nr==113,3,
                     ifelse(logger_nr==111|logger_nr==115,4,pair)))%>%
  mutate(dist=ifelse(pair==3,2.21,
                     ifelse(pair==4,3.90,dist)))
```

```{r message=FALSE, warning=FALSE}
logger_data%>%
  filter(plot=="HC2b")%>%
  group_by(plot,logger_nr,above_below)%>%
  summarise(X=first(X),Y=first(Y))%>%
  ggplot(aes(x=X,y=Y,color=above_below,label=logger_nr))+geom_point()+
  geom_text_repel()+ggtitle("Plot HC2b")+
  scale_y_continuous(breaks = seq(0,13,by=1)) +
  scale_x_continuous(breaks = seq(0,13,by=1)) +
  coord_fixed(ylim=c(0, 5),xlim=c(0, 13))
```

Distance matrix:

```{r message=FALSE, warning=FALSE}
logger_data%>%
  filter(plot=="HC2b")%>%
  group_by(logger_nr)%>%
  summarise(X=first(X),Y=first(Y))%>%
  column_to_rownames(var="logger_nr")%>%
  dist()%>%
  as.matrix()%>%
  # as.data.frame()%>%
  # rownames_to_column(var="logger_nr")%>%
  # pivot_longer(cols=-logger_nr,names_to="logger_nr_2",values_to="distance")
  round(digits=2)
```

Pairs: 122-117 and 119-123

```{r}
logger_data<-logger_data%>%
  mutate(pair=ifelse(logger_nr==122|logger_nr==117,5,
                     ifelse(logger_nr==119|logger_nr==123,6,pair)))%>%
  mutate(dist=ifelse(pair==5,4.47,
                     ifelse(pair==6,2.52,dist)))
```

```{r message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
logger_data%>%
  filter(plot=="HC3")%>%
  group_by(plot,logger_nr,above_below)%>%
  summarise(X=first(X),Y=first(Y))%>%
  ggplot(aes(x=X,y=Y,color=above_below,label=logger_nr))+geom_point()+
  geom_text_repel()+ggtitle("Plot HC3")+
  scale_y_continuous(breaks = seq(0,13,by=1)) +
  scale_x_continuous(breaks = seq(0,13,by=1)) +
  coord_fixed(ylim=c(0, 3),xlim=c(0, 9))
```

Distance matrix:

```{r message=FALSE, warning=FALSE}
logger_data%>%
  filter(plot=="HC3")%>%
  group_by(logger_nr)%>%
  summarise(X=first(X),Y=first(Y))%>%
  column_to_rownames(var="logger_nr")%>%
  dist()%>%
  as.matrix()%>%
  # as.data.frame()%>%
  # rownames_to_column(var="logger_nr")%>%
  # pivot_longer(cols=-logger_nr,names_to="logger_nr_2",values_to="distance")
  round(digits=2)
```

Pairs: 137-127, 138-133 and 136-124

```{r}
logger_data<-logger_data%>%
  mutate(pair=ifelse(logger_nr==137|logger_nr==127,7,
                     ifelse(logger_nr==138|logger_nr==133,8,
                            ifelse(logger_nr==136|logger_nr==124,9,pair))))%>%
  mutate(dist=ifelse(pair==7,0.35,
                     ifelse(pair==8,0.72,
                            ifelse(pair==9,0.22,dist))))
```

```{r message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
logger_data%>%
  filter(plot=="HC4")%>%
  group_by(plot,logger_nr,above_below)%>%
  summarise(X=first(X),Y=first(Y))%>%
  ggplot(aes(x=X,y=Y,color=above_below,label=logger_nr))+geom_point()+
  geom_text_repel()+ggtitle("Plot HC4")+
  scale_y_continuous(breaks = seq(0,20,by=1)) +
  scale_x_continuous(breaks = seq(0,20,by=1)) +
  coord_fixed(ylim=c(0, 10),xlim=c(0, 20))
```

Distance matrix:

```{r message=FALSE, warning=FALSE}
logger_data%>%
  filter(plot=="HC4")%>%
  group_by(logger_nr)%>%
  summarise(X=first(X),Y=first(Y))%>%
  column_to_rownames(var="logger_nr")%>%
  dist()%>%
  as.matrix()%>%
  # as.data.frame()%>%
  # rownames_to_column(var="logger_nr")%>%
  # pivot_longer(cols=-logger_nr,names_to="logger_nr_2",values_to="distance")
  round(digits=2)
```

Pairs: 139-152, 143-153 and 149-154

```{r}
logger_data<-logger_data%>%
  mutate(pair=ifelse(logger_nr==139|logger_nr==152,10,
                     ifelse(logger_nr==143|logger_nr==153,11,
                            ifelse(logger_nr==149|logger_nr==154,12,pair))))%>%
  mutate(dist=ifelse(pair==10,2.63,
                     ifelse(pair==11,0.51,
                            ifelse(pair==12,0.96,dist))))
```

```{r message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
logger_data%>%
  filter(plot=="HC5")%>%
  group_by(plot,logger_nr,above_below)%>%
  summarise(X=first(X),Y=first(Y))%>%
  ggplot(aes(x=X,y=Y,color=above_below,label=logger_nr))+geom_point()+
  geom_text_repel()+ggtitle("Plot HC5")+
  scale_y_continuous(breaks = seq(0,20,by=1)) +
  scale_x_continuous(breaks = seq(0,20,by=1)) +
  coord_fixed(ylim=c(0, 7),xlim=c(0, 13))
```

Distance matrix:

```{r message=FALSE, warning=FALSE}
logger_data%>%
  filter(plot=="HC5")%>%
  group_by(logger_nr)%>%
  summarise(X=first(X),Y=first(Y))%>%
  column_to_rownames(var="logger_nr")%>%
  dist()%>%
  as.matrix()%>%
  # as.data.frame()%>%
  # rownames_to_column(var="logger_nr")%>%
  # pivot_longer(cols=-logger_nr,names_to="logger_nr_2",values_to="distance")
  round(digits=2)
```

Pairs: 165-155, 166-160 and 168-163

```{r}
logger_data<-logger_data%>%
  mutate(pair=ifelse(logger_nr==165|logger_nr==155,13,
                     ifelse(logger_nr==166|logger_nr==160,14,
                            ifelse(logger_nr==168|logger_nr==163,15,pair))))%>%
  mutate(dist=ifelse(pair==13,1.47,
                     ifelse(pair==14,1.08,
                            ifelse(pair==15,1.35,dist))))
```

```{r message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
logger_data%>%
  filter(plot=="HC6")%>%
  group_by(plot,logger_nr,above_below)%>%
  summarise(X=first(X),Y=first(Y))%>%
  ggplot(aes(x=X,y=Y,color=above_below,label=logger_nr))+geom_point()+
  geom_text_repel()+ggtitle("Plot HC6")+
  scale_y_continuous(breaks = seq(0,22,by=1)) +
  scale_x_continuous(breaks = seq(0,22,by=1)) +
  coord_fixed(ylim=c(0, 10),xlim=c(0, 20))
```

Distance matrix:

```{r message=FALSE, warning=FALSE}
logger_data%>%
  filter(plot=="HC6")%>%
  group_by(logger_nr)%>%
  summarise(X=first(X),Y=first(Y))%>%
  column_to_rownames(var="logger_nr")%>%
  dist()%>%
  as.matrix()%>%
  # as.data.frame()%>%
  # rownames_to_column(var="logger_nr")%>%
  # pivot_longer(cols=-logger_nr,names_to="logger_nr_2",values_to="distance")
  round(digits=2)
```

Pairs: 205-169 and 209-202

```{r}
logger_data<-logger_data%>%
  mutate(pair=ifelse(logger_nr==205|logger_nr==169,16,
                     ifelse(logger_nr==209|logger_nr==202,17,pair)))%>%
  mutate(dist=ifelse(pair==16,1.30,
                     ifelse(pair==17,4.87,dist)))
```

```{r message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
logger_data%>%
  filter(plot=="HC10")%>%
  group_by(plot,logger_nr,above_below)%>%
  summarise(X=first(X),Y=first(Y))%>%
  ggplot(aes(x=X,y=Y,color=above_below,label=logger_nr))+geom_point()+
  geom_text_repel()+ggtitle("Plot HC10")+
  scale_y_continuous(breaks = seq(0,22,by=1)) +
  scale_x_continuous(breaks = seq(0,22,by=1)) +
  coord_fixed(ylim=c(0, 7),xlim=c(0, 7))
```

Distance matrix:

```{r message=FALSE, warning=FALSE}
logger_data%>%
  filter(plot=="HC10")%>%
  group_by(logger_nr)%>%
  summarise(X=first(X),Y=first(Y))%>%
  column_to_rownames(var="logger_nr")%>%
  dist()%>%
  as.matrix()%>%
  # as.data.frame()%>%
  # rownames_to_column(var="logger_nr")%>%
  # pivot_longer(cols=-logger_nr,names_to="logger_nr_2",values_to="distance")
  round(digits=2)
```

Pairs: 57-3, 71-20 and 58-28

```{r}
logger_data<-logger_data%>%
  mutate(pair=ifelse(logger_nr==57|logger_nr==3,18,
                     ifelse(logger_nr==71|logger_nr==20,19,
                            ifelse(logger_nr==58|logger_nr==28,20,pair))))%>%
  mutate(dist=ifelse(pair==18,1.27,
                     ifelse(pair==19,1.01,
                            ifelse(pair==20,1.34,dist))))
```

```{r message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
logger_data%>%
  filter(plot=="HC7")%>%
  group_by(plot,logger_nr,above_below)%>%
  summarise(X=first(X),Y=first(Y))%>%
  ggplot(aes(x=X,y=Y,color=above_below,label=logger_nr))+geom_point()+
  geom_text_repel()+ggtitle("Plot HC7")+
  scale_y_continuous(breaks = seq(0,22,by=1)) +
  scale_x_continuous(breaks = seq(0,22,by=1)) +
  coord_fixed(ylim=c(0, 4),xlim=c(0, 13))
```

Distance matrix:

```{r message=FALSE, warning=FALSE}
logger_data%>%
  filter(plot=="HC7")%>%
  group_by(logger_nr)%>%
  summarise(X=first(X),Y=first(Y))%>%
  column_to_rownames(var="logger_nr")%>%
  dist()%>%
  as.matrix()%>%
  # as.data.frame()%>%
  # rownames_to_column(var="logger_nr")%>%
  # pivot_longer(cols=-logger_nr,names_to="logger_nr_2",values_to="distance")
  round(digits=2)
```

Pairs: 220-212, 221-215 and 222-218

```{r}
logger_data<-logger_data%>%
  mutate(pair=ifelse(logger_nr==220|logger_nr==212,21,
                     ifelse(logger_nr==221|logger_nr==215,22,
                            ifelse(logger_nr==222|logger_nr==218,23,pair))))%>%
  mutate(dist=ifelse(pair==21,1.91,
                     ifelse(pair==22,1.41,
                            ifelse(pair==23,1.50,dist))))
```

```{r message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
logger_data%>%
  filter(plot=="HC8")%>%
  group_by(plot,logger_nr,above_below)%>%
  summarise(X=first(X),Y=first(Y))%>%
  ggplot(aes(x=X,y=Y,color=above_below,label=logger_nr))+geom_point()+
  geom_text_repel()+ggtitle("Plot HC8")+
  scale_y_continuous(breaks = seq(0,22,by=1)) +
  scale_x_continuous(breaks = seq(0,22,by=1)) +
  coord_fixed(ylim=c(0, 6),xlim=c(0, 10))
```

Distance matrix:

```{r message=FALSE, warning=FALSE}
logger_data%>%
  filter(plot=="HC8")%>%
  group_by(logger_nr)%>%
  summarise(X=first(X),Y=first(Y))%>%
  column_to_rownames(var="logger_nr")%>%
  dist()%>%
  as.matrix()%>%
  # as.data.frame()%>%
  # rownames_to_column(var="logger_nr")%>%
  # pivot_longer(cols=-logger_nr,names_to="logger_nr_2",values_to="distance")
  round(digits=2)
```

Pairs: 233-224, 234-232 and 237-230

```{r}
logger_data<-logger_data%>%
  mutate(pair=ifelse(logger_nr==233|logger_nr==224,24,
                     ifelse(logger_nr==234|logger_nr==232,25,
                            ifelse(logger_nr==237|logger_nr==230,26,pair))))%>%
  mutate(dist=ifelse(pair==24,2.60,
                     ifelse(pair==25,1.33,
                            ifelse(pair==26,1.10,dist))))
```

```{r message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
logger_data%>%
  filter(plot=="HC09")%>%
  group_by(plot,logger_nr,above_below)%>%
  summarise(X=first(X),Y=first(Y))%>%
  ggplot(aes(x=X,y=Y,color=above_below,label=logger_nr))+geom_point()+
  geom_text_repel()+ggtitle("Plot HC09")+
  scale_y_continuous(breaks = seq(0,22,by=1)) +
  scale_x_continuous(breaks = seq(0,22,by=1)) +
  coord_fixed(ylim=c(0, 6),xlim=c(0, 20))
```

Distance matrix:

```{r message=FALSE, warning=FALSE}
logger_data%>%
  filter(plot=="HC09")%>%
  group_by(logger_nr)%>%
  summarise(X=first(X),Y=first(Y))%>%
  column_to_rownames(var="logger_nr")%>%
  dist()%>%
  as.matrix()%>%
  # as.data.frame()%>%
  # rownames_to_column(var="logger_nr")%>%
  # pivot_longer(cols=-logger_nr,names_to="logger_nr_2",values_to="distance")
  round(digits=2)
```

Pairs: 249-238, 251-243 and 252-247

```{r}
logger_data<-logger_data%>%
  mutate(pair=ifelse(logger_nr==249|logger_nr==238,27,
                     ifelse(logger_nr==251|logger_nr==243,28,
                            ifelse(logger_nr==252|logger_nr==247,29,pair))))%>%
  mutate(dist=ifelse(pair==27,2.24,
                     ifelse(pair==28,2.55,
                            ifelse(pair==29,1.81,dist))))
```

```{r message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
logger_data%>%
  filter(plot=="H10")%>%
  group_by(plot,logger_nr,above_below)%>%
  summarise(X=first(X),Y=first(Y))%>%
  ggplot(aes(x=X,y=Y,color=above_below,label=logger_nr))+geom_point()+
  geom_text_repel()+ggtitle("Plot H10")+
  scale_y_continuous(breaks = seq(-5,30,by=1)) +
  scale_x_continuous(breaks = seq(0,30,by=1)) +
  coord_fixed(ylim=c(-5, 8),xlim=c(0, 30))
```

Distance matrix:

```{r message=FALSE, warning=FALSE}
logger_data%>%
  filter(plot=="H10")%>%
  group_by(logger_nr)%>%
  summarise(X=first(X),Y=first(Y))%>%
  column_to_rownames(var="logger_nr")%>%
  dist()%>%
  as.matrix()%>%
  # as.data.frame()%>%
  # rownames_to_column(var="logger_nr")%>%
  # pivot_longer(cols=-logger_nr,names_to="logger_nr_2",values_to="distance")
  round(digits=2)
```

Pairs: 261-258, 262-255 (same plant/location), 263-256 (same plant/location)

```{r}
logger_data<-logger_data%>%
  mutate(pair=ifelse(logger_nr==261|logger_nr==258,30,
                     ifelse(logger_nr==262|logger_nr==255,31,
                            ifelse(logger_nr==263|logger_nr==256,32,pair))))%>%
  mutate(dist=ifelse(pair==30,13.32,
                     ifelse(pair==31,0,
                            ifelse(pair==32,0,dist))))
```

```{r message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
logger_data%>%
  filter(plot=="H01")%>%
  group_by(plot,logger_nr,above_below)%>%
  summarise(X=first(X),Y=first(Y))%>%
  ggplot(aes(x=X,y=Y,color=above_below,label=logger_nr))+geom_point()+
  geom_text_repel()+ggtitle("Plot H01")+
  scale_y_continuous(breaks = seq(0,22,by=1)) +
  scale_x_continuous(breaks = seq(0,22,by=1)) +
  coord_fixed(ylim=c(0, 8),xlim=c(0, 20))
```

Distance matrix:

```{r message=FALSE, warning=FALSE}
logger_data%>%
  filter(plot=="H01")%>%
  group_by(logger_nr)%>%
  summarise(X=first(X),Y=first(Y))%>%
  column_to_rownames(var="logger_nr")%>%
  dist()%>%
  as.matrix()%>%
  # as.data.frame()%>%
  # rownames_to_column(var="logger_nr")%>%
  # pivot_longer(cols=-logger_nr,names_to="logger_nr_2",values_to="distance")
  round(digits=2)
```

Pairs: 273-266, 274-267 and 275-270

```{r}
logger_data<-logger_data%>%
  mutate(pair=ifelse(logger_nr==273|logger_nr==266,33,
                     ifelse(logger_nr==274|logger_nr==267,34,
                            ifelse(logger_nr==275|logger_nr==270,35,pair))))%>%
  mutate(dist=ifelse(pair==33,3.09,
                     ifelse(pair==34,1.84,
                            ifelse(pair==35,2.15,dist))))
```

```{r message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
logger_data%>%
  filter(plot=="H02")%>%
  group_by(plot,logger_nr,above_below)%>%
  summarise(X=first(X),Y=first(Y))%>%
  ggplot(aes(x=X,y=Y,color=above_below,label=logger_nr))+geom_point()+
  geom_text_repel()+ggtitle("Plot H02")+
  scale_y_continuous(breaks = seq(0,22,by=1)) +
  scale_x_continuous(breaks = seq(0,22,by=1)) +
  coord_fixed(ylim=c(0, 10),xlim=c(0, 15))
```

Distance matrix:

```{r message=FALSE, warning=FALSE}
logger_data%>%
  filter(plot=="H02")%>%
  group_by(logger_nr)%>%
  summarise(X=first(X),Y=first(Y))%>%
  column_to_rownames(var="logger_nr")%>%
  dist()%>%
  as.matrix()%>%
  # as.data.frame()%>%
  # rownames_to_column(var="logger_nr")%>%
  # pivot_longer(cols=-logger_nr,names_to="logger_nr_2",values_to="distance")
  round(digits=2)
```

Pairs: 287-279

```{r}
logger_data<-logger_data%>%
  mutate(pair=ifelse(logger_nr==287|logger_nr==279,36,pair))%>%
  mutate(dist=ifelse(pair==36,9.06,dist))
```

```{r message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
logger_data%>%
  filter(plot=="H03")%>%
  group_by(plot,logger_nr,above_below)%>%
  summarise(X=first(X),Y=first(Y))%>%
  ggplot(aes(x=X,y=Y,color=above_below,label=logger_nr))+geom_point()+
  geom_text_repel()+ggtitle("Plot H03")+
  scale_y_continuous(breaks = seq(0,22,by=1)) +
  scale_x_continuous(breaks = seq(0,22,by=1)) +
  coord_fixed(ylim=c(0, 7),xlim=c(0, 12))
```

Distance matrix:

```{r message=FALSE, warning=FALSE}
logger_data%>%
  filter(plot=="H03")%>%
  group_by(logger_nr)%>%
  summarise(X=first(X),Y=first(Y))%>%
  column_to_rownames(var="logger_nr")%>%
  dist()%>%
  as.matrix()%>%
  # as.data.frame()%>%
  # rownames_to_column(var="logger_nr")%>%
  # pivot_longer(cols=-logger_nr,names_to="logger_nr_2",values_to="distance")
  round(digits=2)
```

Pairs: 296-288, 297-289 and 298-290 (same locations/plants)

```{r}
logger_data<-logger_data%>%
  mutate(pair=ifelse(logger_nr==296|logger_nr==288,37,
                     ifelse(logger_nr==297|logger_nr==289,38,
                            ifelse(logger_nr==298|logger_nr==290,39,pair))))%>%
  mutate(dist=ifelse(pair==37,0,
                     ifelse(pair==38,0,
                            ifelse(pair==39,0,dist))))
```

```{r message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
logger_data%>%
  filter(plot=="H04")%>%
  group_by(plot,logger_nr,above_below)%>%
  summarise(X=first(X),Y=first(Y))%>%
  ggplot(aes(x=X,y=Y,color=above_below,label=logger_nr))+geom_point()+
  geom_text_repel()+ggtitle("Plot H04")+
  scale_y_continuous(breaks = seq(-2,30,by=1)) +
  scale_x_continuous(breaks = seq(0,30,by=1)) +
  coord_fixed(ylim=c(-2, 7),xlim=c(0, 28))
```

Distance matrix:

```{r message=FALSE, warning=FALSE}
logger_data%>%
  filter(plot=="H04")%>%
  group_by(logger_nr)%>%
  summarise(X=first(X),Y=first(Y))%>%
  column_to_rownames(var="logger_nr")%>%
  dist()%>%
  as.matrix()%>%
  # as.data.frame()%>%
  # rownames_to_column(var="logger_nr")%>%
  # pivot_longer(cols=-logger_nr,names_to="logger_nr_2",values_to="distance")
  round(digits=2)
```

Pairs: 306-300, 307-301 and 308-302

```{r}
logger_data<-logger_data%>%
  mutate(pair=ifelse(logger_nr==306|logger_nr==300,40,
                     ifelse(logger_nr==307|logger_nr==301,41,
                            ifelse(logger_nr==308|logger_nr==302,42,pair))))%>%
  mutate(dist=ifelse(pair==40,1.67,
                     ifelse(pair==41,4.04,
                            ifelse(pair==42,1.19,dist))))
```

```{r message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
logger_data%>%
  filter(plot=="H05")%>%
  group_by(plot,logger_nr,above_below)%>%
  summarise(X=first(X),Y=first(Y))%>%
  ggplot(aes(x=X,y=Y,color=above_below,label=logger_nr))+geom_point()+
  geom_text_repel()+ggtitle("Plot H05")+
  scale_y_continuous(breaks = seq(0,30,by=1)) +
  scale_x_continuous(breaks = seq(0,30,by=1)) +
  coord_fixed(ylim=c(0, 6),xlim=c(0, 20))
```

Distance matrix:

```{r message=FALSE, warning=FALSE}
logger_data%>%
  filter(plot=="H05")%>%
  group_by(logger_nr)%>%
  summarise(X=first(X),Y=first(Y))%>%
  column_to_rownames(var="logger_nr")%>%
  dist()%>%
  as.matrix()%>%
  # as.data.frame()%>%
  # rownames_to_column(var="logger_nr")%>%
  # pivot_longer(cols=-logger_nr,names_to="logger_nr_2",values_to="distance")
  round(digits=2)
```

Pairs: 319-310, 320-312 and 321-317

```{r}
logger_data<-logger_data%>%
  mutate(pair=ifelse(logger_nr==319|logger_nr==310,43,
                     ifelse(logger_nr==320|logger_nr==312,44,
                            ifelse(logger_nr==321|logger_nr==317,45,pair))))%>%
  mutate(dist=ifelse(pair==43,1.46,
                     ifelse(pair==44,1.32,
                            ifelse(pair==45,0.96,dist))))
```

```{r message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
logger_data%>%
  filter(plot=="H06")%>%
  group_by(plot,logger_nr,above_below)%>%
  summarise(X=first(X),Y=first(Y))%>%
  ggplot(aes(x=X,y=Y,color=above_below,label=logger_nr))+geom_point()+
  geom_text_repel()+ggtitle("Plot H06")+
  scale_y_continuous(breaks = seq(0,30,by=1)) +
  scale_x_continuous(breaks = seq(0,30,by=1)) +
  coord_fixed(ylim=c(0, 6),xlim=c(0, 28))
```

Distance matrix:

```{r message=FALSE, warning=FALSE}
logger_data%>%
  filter(plot=="H06")%>%
  group_by(logger_nr)%>%
  summarise(X=first(X),Y=first(Y))%>%
  column_to_rownames(var="logger_nr")%>%
  dist()%>%
  as.matrix()%>%
  # as.data.frame()%>%
  # rownames_to_column(var="logger_nr")%>%
  # pivot_longer(cols=-logger_nr,names_to="logger_nr_2",values_to="distance")
  round(digits=2)
```

Pairs: 331-323, 332-324 and 333-329

```{r}
logger_data<-logger_data%>%
  mutate(pair=ifelse(logger_nr==331|logger_nr==323,46,
                     ifelse(logger_nr==332|logger_nr==324,47,
                            ifelse(logger_nr==333|logger_nr==329,48,pair))))%>%
  mutate(dist=ifelse(pair==46,2.25,
                     ifelse(pair==47,2.39,
                            ifelse(pair==48,1.29,dist))))
```

```{r message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
logger_data%>%
  filter(plot=="H07")%>%
  group_by(plot,logger_nr,above_below)%>%
  summarise(X=first(X),Y=first(Y))%>%
  ggplot(aes(x=X,y=Y,color=above_below,label=logger_nr))+geom_point()+
  geom_text_repel()+ggtitle("Plot H07")+
  scale_y_continuous(breaks = seq(0,30,by=1)) +
  scale_x_continuous(breaks = seq(0,30,by=1)) +
  coord_fixed(ylim=c(0, 3),xlim=c(0, 20))
```

Distance matrix:

```{r message=FALSE, warning=FALSE}
logger_data%>%
  filter(plot=="H07")%>%
  group_by(logger_nr)%>%
  summarise(X=first(X),Y=first(Y))%>%
  column_to_rownames(var="logger_nr")%>%
  dist()%>%
  as.matrix()%>%
  # as.data.frame()%>%
  # rownames_to_column(var="logger_nr")%>%
  # pivot_longer(cols=-logger_nr,names_to="logger_nr_2",values_to="distance")
  round(digits=2)
```

Pairs: 344-334, 345-336 and 346-339

```{r}
logger_data<-logger_data%>%
  mutate(pair=ifelse(logger_nr==344|logger_nr==334,49,
                     ifelse(logger_nr==345|logger_nr==336,50,
                            ifelse(logger_nr==346|logger_nr==339,51,pair))))%>%
  mutate(dist=ifelse(pair==49,1.11,
                     ifelse(pair==50,1.48,
                            ifelse(pair==51,1.52,dist))))
```

```{r message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
logger_data%>%
  filter(plot=="H08")%>%
  group_by(plot,logger_nr,above_below)%>%
  summarise(X=first(X),Y=first(Y))%>%
  ggplot(aes(x=X,y=Y,color=above_below,label=logger_nr))+geom_point()+
  geom_text_repel()+ggtitle("Plot H08")+
  scale_y_continuous(breaks = seq(0,30,by=1)) +
  scale_x_continuous(breaks = seq(0,30,by=1)) +
  coord_fixed(ylim=c(-1, 5),xlim=c(0, 21))
```

Distance matrix:

```{r message=FALSE, warning=FALSE}
logger_data%>%
  filter(plot=="H08")%>%
  group_by(logger_nr)%>%
  summarise(X=first(X),Y=first(Y))%>%
  column_to_rownames(var="logger_nr")%>%
  dist()%>%
  as.matrix()%>%
  # as.data.frame()%>%
  # rownames_to_column(var="logger_nr")%>%
  # pivot_longer(cols=-logger_nr,names_to="logger_nr_2",values_to="distance")
  round(digits=2)
```

Pairs: 355-347, 356-349 and 357-352

```{r}
logger_data<-logger_data%>%
  mutate(pair=ifelse(logger_nr==355|logger_nr==347,52,
                     ifelse(logger_nr==356|logger_nr==349,53,
                            ifelse(logger_nr==357|logger_nr==352,54,pair))))%>%
  mutate(dist=ifelse(pair==52,1.24,
                     ifelse(pair==53,4.59,
                            ifelse(pair==54,3.54,dist))))
```

```{r message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
logger_data%>%
  filter(plot=="H09")%>%
  group_by(plot,logger_nr,above_below)%>%
  summarise(X=first(X),Y=first(Y))%>%
  ggplot(aes(x=X,y=Y,color=above_below,label=logger_nr))+geom_point()+
  geom_text_repel()+ggtitle("Plot H09")+
  scale_y_continuous(breaks = seq(0,30,by=1)) +
  scale_x_continuous(breaks = seq(0,30,by=1)) +
  coord_fixed(ylim=c(0, 4),xlim=c(0, 20))
```

Distance matrix:

```{r message=FALSE, warning=FALSE}
logger_data%>%
  filter(plot=="H09")%>%
  group_by(logger_nr)%>%
  summarise(X=first(X),Y=first(Y))%>%
  column_to_rownames(var="logger_nr")%>%
  dist()%>%
  as.matrix()%>%
  # as.data.frame()%>%
  # rownames_to_column(var="logger_nr")%>%
  # pivot_longer(cols=-logger_nr,names_to="logger_nr_2",values_to="distance")
  round(digits=2)
```

Pairs: 369-362, 370-366 and 371-367

```{r}
logger_data<-logger_data%>%
  mutate(pair=ifelse(logger_nr==369|logger_nr==362,55,
                     ifelse(logger_nr==370|logger_nr==366,56,
                            ifelse(logger_nr==371|logger_nr==367,57,pair))))%>%
  mutate(dist=ifelse(pair==55,2.55,
                     ifelse(pair==56,4.68,
                            ifelse(pair==57,1.19,dist))))
```

Plot H11 not included because there are only aboveground loggers

```{r message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
logger_data%>%
  filter(plot=="H13")%>%
  group_by(plot,logger_nr,above_below)%>%
  summarise(X=first(X),Y=first(Y))%>%
  ggplot(aes(x=X,y=Y,color=above_below,label=logger_nr))+geom_point()+
  geom_text_repel()+ggtitle("Plot H13")+
  scale_y_continuous(breaks = seq(0,30,by=1)) +
  scale_x_continuous(breaks = seq(0,30,by=1)) +
  coord_fixed(ylim=c(-1, 3),xlim=c(0, 11))
```

Distance matrix:

```{r message=FALSE, warning=FALSE}
logger_data%>%
  filter(plot=="H13")%>%
  group_by(logger_nr)%>%
  summarise(X=first(X),Y=first(Y))%>%
  column_to_rownames(var="logger_nr")%>%
  dist()%>%
  as.matrix()%>%
  # as.data.frame()%>%
  # rownames_to_column(var="logger_nr")%>%
  # pivot_longer(cols=-logger_nr,names_to="logger_nr_2",values_to="distance")
  round(digits=2)
```

Pairs: 395-386 (same location/plant) and 396-391

```{r}
logger_data<-logger_data%>%
  mutate(pair=ifelse(logger_nr==395|logger_nr==386,58,
                     ifelse(logger_nr==396|logger_nr==391,59,pair)))%>%
  mutate(dist=ifelse(pair==58,0,
                     ifelse(pair==59,5.16,dist)))
```

```{r message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
logger_data%>%
  filter(plot=="H14")%>%
  group_by(plot,logger_nr,above_below)%>%
  summarise(X=first(X),Y=first(Y))%>%
  ggplot(aes(x=X,y=Y,color=above_below,label=logger_nr))+geom_point()+
  geom_text_repel()+ggtitle("Plot H14")+
  scale_y_continuous(breaks = seq(0,30,by=1)) +
  scale_x_continuous(breaks = seq(0,30,by=1)) +
  coord_fixed(ylim=c(0,7),xlim=c(0, 9))
```

Distance matrix:

```{r message=FALSE, warning=FALSE}
logger_data%>%
  filter(plot=="H14")%>%
  group_by(logger_nr)%>%
  summarise(X=first(X),Y=first(Y))%>%
  column_to_rownames(var="logger_nr")%>%
  dist()%>%
  as.matrix()%>%
  # as.data.frame()%>%
  # rownames_to_column(var="logger_nr")%>%
  # pivot_longer(cols=-logger_nr,names_to="logger_nr_2",values_to="distance")
  round(digits=2)
```

Pairs: 411-406, 408-397 and 415-407

```{r}
logger_data<-logger_data%>%
  mutate(pair=ifelse(logger_nr==411|logger_nr==406,60,
                     ifelse(logger_nr==408|logger_nr==397,61,
                            ifelse(logger_nr==415|logger_nr==407,62,pair))))%>%
  mutate(dist=ifelse(pair==60,1.50,
                     ifelse(pair==61,0,
                            ifelse(pair==62,3.45,dist))))
```

Plot H11 not included because there are only aboveground loggers

```{r message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
logger_data%>%
  filter(plot=="H16")%>%
  group_by(plot,logger_nr,above_below)%>%
  summarise(X=first(X),Y=first(Y))%>%
  ggplot(aes(x=X,y=Y,color=above_below,label=logger_nr))+geom_point()+
  geom_text_repel()+ggtitle("Plot H16")+
  scale_y_continuous(breaks = seq(0,30,by=1)) +
  scale_x_continuous(breaks = seq(0,30,by=1)) +
  coord_fixed(ylim=c(0, 12),xlim=c(0, 5))
```

Distance matrix:

```{r message=FALSE, warning=FALSE}
logger_data%>%
  filter(plot=="H16")%>%
  group_by(logger_nr)%>%
  summarise(X=first(X),Y=first(Y))%>%
  column_to_rownames(var="logger_nr")%>%
  dist()%>%
  as.matrix()%>%
  # as.data.frame()%>%
  # rownames_to_column(var="logger_nr")%>%
  # pivot_longer(cols=-logger_nr,names_to="logger_nr_2",values_to="distance")
  round(digits=2)
```

Pairs: 458-559

```{r}
logger_data<-logger_data%>%
  mutate(pair=ifelse(logger_nr==458|logger_nr==559,63,pair))%>%
  mutate(dist=ifelse(pair==63,0,dist))
```

```{r message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
logger_data%>%
  filter(plot=="H17")%>%
  group_by(plot,logger_nr,above_below)%>%
  summarise(X=first(X),Y=first(Y))%>%
  ggplot(aes(x=X,y=Y,color=above_below,label=logger_nr))+geom_point()+
  geom_text_repel()+ggtitle("Plot H17")+
  scale_y_continuous(breaks = seq(0,30,by=1)) +
  scale_x_continuous(breaks = seq(0,30,by=1)) +
  coord_fixed(ylim=c(0, 8),xlim=c(0, 9))
```

Distance matrix:

```{r message=FALSE, warning=FALSE}
logger_data%>%
  filter(plot=="H17")%>%
  group_by(logger_nr)%>%
  summarise(X=first(X),Y=first(Y))%>%
  column_to_rownames(var="logger_nr")%>%
  dist()%>%
  as.matrix()%>%
  # as.data.frame()%>%
  # rownames_to_column(var="logger_nr")%>%
  # pivot_longer(cols=-logger_nr,names_to="logger_nr_2",values_to="distance")
  round(digits=2)
```

Pairs: 654-585, 655-588 and 656-599 (all same locations / plants)

```{r}
logger_data<-logger_data%>%
  mutate(pair=ifelse(logger_nr==654|logger_nr==585,64,
                     ifelse(logger_nr==655|logger_nr==588,65,
                            ifelse(logger_nr==656|logger_nr==599,66,pair))))%>%
  mutate(dist=ifelse(pair==64,0,
                     ifelse(pair==65,0,
                            ifelse(pair==66,0,dist))))
```

```{r message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
logger_data%>%
  filter(plot=="H18a")%>%
  group_by(plot,logger_nr,above_below)%>%
  summarise(X=first(X),Y=first(Y))%>%
  ggplot(aes(x=X,y=Y,color=above_below,label=logger_nr))+geom_point()+
  geom_text_repel()+ggtitle("Plot H18a")+
  scale_y_continuous(breaks = seq(0,30,by=1)) +
  scale_x_continuous(breaks = seq(0,30,by=1)) +
  coord_fixed(ylim=c(0, 3),xlim=c(0, 25))
```

Distance matrix:

```{r message=FALSE, warning=FALSE}
logger_data%>%
  filter(plot=="H18a")%>%
  group_by(logger_nr)%>%
  summarise(X=first(X),Y=first(Y))%>%
  column_to_rownames(var="logger_nr")%>%
  dist()%>%
  as.matrix()%>%
  # as.data.frame()%>%
  # rownames_to_column(var="logger_nr")%>%
  # pivot_longer(cols=-logger_nr,names_to="logger_nr_2",values_to="distance")
  round(digits=2)
```

Pairs: 663-659 and 664-662 (all same locations / plants)

```{r}
logger_data<-logger_data%>%
  mutate(pair=ifelse(logger_nr==663|logger_nr==659,67,
                     ifelse(logger_nr==664|logger_nr==662,68,pair)))%>%
  mutate(dist=ifelse(pair==67,0,
                     ifelse(pair==68,0,dist)))
```

```{r message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
logger_data%>%
  filter(plot=="H18b")%>%
  group_by(plot,logger_nr,above_below)%>%
  summarise(X=first(X),Y=first(Y))%>%
  ggplot(aes(x=X,y=Y,color=above_below,label=logger_nr))+geom_point()+
  geom_text_repel()+ggtitle("Plot H18b")+
  scale_y_continuous(breaks = seq(0,30,by=1)) +
  scale_x_continuous(breaks = seq(0,30,by=1)) +
  coord_fixed(ylim=c(0, 3),xlim=c(0, 3))
```

Distance matrix:

```{r message=FALSE, warning=FALSE}
logger_data%>%
  filter(plot=="H18b")%>%
  group_by(logger_nr)%>%
  summarise(X=first(X),Y=first(Y))%>%
  column_to_rownames(var="logger_nr")%>%
  dist()%>%
  as.matrix()%>%
  # as.data.frame()%>%
  # rownames_to_column(var="logger_nr")%>%
  # pivot_longer(cols=-logger_nr,names_to="logger_nr_2",values_to="distance")
  round(digits=2)
```

Pairs: 671-668 (same location / plant) and 670-666

```{r}
logger_data<-logger_data%>%
  mutate(pair=ifelse(logger_nr==671|logger_nr==668,69,
                     ifelse(logger_nr==670|logger_nr==666,70,pair)))%>%
  mutate(dist=ifelse(pair==69,0,
                     ifelse(pair==70,1.77,dist)))
```

```{r message=FALSE, warning=FALSE, fig.height=5, fig.width=6}
logger_data%>%
  filter(plot=="H19")%>%
  group_by(plot,logger_nr,above_below)%>%
  summarise(X=first(X),Y=first(Y))%>%
  ggplot(aes(x=X,y=Y,color=above_below,label=logger_nr))+geom_point()+
  geom_text_repel()+ggtitle("Plot H19")+
  scale_y_continuous(breaks = seq(0,30,by=1)) +
  scale_x_continuous(breaks = seq(0,30,by=1)) +
  coord_fixed(ylim=c(0, 12),xlim=c(0, 4))
```

Distance matrix:

```{r message=FALSE, warning=FALSE}
logger_data%>%
  filter(plot=="H19")%>%
  group_by(logger_nr)%>%
  summarise(X=first(X),Y=first(Y))%>%
  column_to_rownames(var="logger_nr")%>%
  dist()%>%
  as.matrix()%>%
  # as.data.frame()%>%
  # rownames_to_column(var="logger_nr")%>%
  # pivot_longer(cols=-logger_nr,names_to="logger_nr_2",values_to="distance")
  round(digits=2)
```

Pairs: 683-675, 682-674, 681-673 (all same locations / plants)

```{r}
logger_data<-logger_data%>%
  mutate(pair=ifelse(logger_nr==683|logger_nr==675,71,
                     ifelse(logger_nr==682|logger_nr==674,72,
                            ifelse(logger_nr==681|logger_nr==673,73,pair))))%>%
  mutate(dist=ifelse(pair==71,0,
                     ifelse(pair==72,0,
                            ifelse(pair==73,0,dist))))
```

```{r}
logger_data_pairs<-subset(logger_data,!is.na(pair))
```

# Write logger_data and logger_data_pairs to csv files

```{r include=FALSE}
write_csv(logger_data, file = "data/edited/logger_data.csv")
write_csv(logger_data_pairs, file = "data/edited/logger_data_pairs.csv")
```

```{r include=FALSE}
sessionInfo()
```




