---
title: "Maladaptive plastic responses of flowering time to geothermal heating (Cerastium 2)"
subtitle: "Analyses with logger data"
author : "Alicia Valdés"
output:
  pdf_document:
    toc: yes
    toc_depth: 4
  html_notebook:
    toc: yes
    toc_depth: '4'
    latex_engine: xelatex
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r load packages, include=FALSE}
library(tidyverse)
library(ggthemes)
library(knitr)
library(gridExtra)
library(DHARMa)
library(RColorBrewer)
library(broom)
library(ggpubr)
library(jtools)
library(kableExtra)
library(ggeffects)
library(MuMIn)
library(MASS)
library(segmented)
library(effects)
library(purrr)
library(lubridate)
library(ggforce)
library(car)
library(ggrepel)
```

```{r Define ggplot themes and palettes, include=FALSE}
my_theme <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(legend.position="none")+theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
my_theme_legend <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
myPalette <- colorRampPalette(brewer.pal(11, "YlOrRd"))
```

```{r, include=FALSE}
set_summ_defaults(digits = 3)
```

# Read data

```{r}
logger_data<-read_csv("data/edited/logger_data.csv")
logger_data_pairs<-read_csv("data/edited/logger_data_pairs.csv")
```

# Plot with all logger data, one line per logger id

```{r fig.height=5, fig.width=8}
ggplot(logger_data,aes(x=datetime,y=temp,group=logger_nr))+
  geom_line(size=0.1)
```

# Q1: Are instantaneous measures of soil temperature representative for the conditions during the entire spring/growing season?

Correlations logger temperature - instant temperature

## May

For each logger\_nr, get mean temperature during May 2017 and compare with temp\_term (which was measured with a thermometer at 10 cm depth on May 2017):

```{r message=FALSE, warning=FALSE, fig.height=3, fig.width=5}
maytemps<-logger_data%>%
  mutate(month = month(datetime)) %>%
  filter(month==5)%>%
  group_by(logger_nr) %>%
  summarize(meanmay_logger=mean(temp),temp_term=mean(temp_term),
            above_below=first(above_below))
ggplot(maytemps,aes(x=temp_term,y=meanmay_logger,color=above_below))+
  geom_point(size=0.5)+geom_smooth(method="lm",size=0.5)+
  geom_abline(lty="dashed",alpha=0.8)+
  xlab("Soil temperature (10 cm depth) measured in May")+
  ylab("Mean temperature of May\nfrom logger data")+
  scale_color_manual(name="Location",values=c("#F8766D", "#00BFC4"),
                     labels = c("Above, slope=0.04","Below, slope=0.85"))
summ(lm(meanmay_logger~temp_term,subset(maytemps,above_below=="A")))
summ(lm(meanmay_logger~temp_term,subset(maytemps,above_below=="B")))
```

Correlation mean temperature of may from logger data and soil temperature measured in may with thermometer:

```{r}
with(maytemps,cor(meanmay_logger,temp_term))
```

Correlation mean temperature of may from logger data (only belowground loggers) and soil temperature measured in may with thermometer:

```{r}
with(subset(maytemps,above_below=="B"),cor(meanmay_logger,temp_term))
```

Correlation mean temperature of may from logger data (only aboveground loggers) and soil temperature measured in may with thermometer:

```{r}
with(subset(maytemps,above_below=="A"),cor(meanmay_logger,temp_term))
```

## Whole period

For each logger\_nr, get mean temperature during the whole period available and compare with temp\_term (which was measured with a thermometer at 10 cm depth on May 2017):

```{r message=FALSE, warning=FALSE, fig.height=3, fig.width=5}
alltemps<-logger_data%>%
  group_by(logger_nr) %>%
  summarize(mean_logger=mean(temp),temp_term=mean(temp_term),
            above_below=first(above_below))
ggplot(alltemps,aes(x=temp_term,y=mean_logger,color=above_below))+
  geom_point(size=0.5)+geom_smooth(method="lm",size=0.5)+
  geom_abline(lty="dashed",alpha=0.8)+
  xlab("Soil temperature (10 cm depth) measured in May")+
  ylab("Mean temperature\nfrom logger data (whole period)")+
  scale_color_manual(name="Location",values=c("#F8766D", "#00BFC4"),
                     labels = c("Above, slope=0.07","Below, slope=0.83"))
summ(lm(mean_logger~temp_term,subset(alltemps,above_below=="A")))
summ(lm(mean_logger~temp_term,subset(alltemps,above_below=="B")))
```

Correlation mean temperature from logger data and soil temperature measured in may with thermometer:

```{r}
with(alltemps,cor(mean_logger,temp_term))
```

Correlation mean temperature from logger data (only belowground loggers) and soil temperature measured in may with thermometer:

```{r}
with(subset(alltemps,above_below=="B"),cor(mean_logger,temp_term))
```

Correlation mean temperature from logger data (only aboveground loggers) and soil temperature measured in may with thermometer:

```{r}
with(subset(alltemps,above_below=="A"),cor(mean_logger,temp_term))
```

The correlation values seem to indicate that the temperature measured with a thermometer represents quite well longer-term conditions

# Q2: Do differences between soil and air temperatures change with soil temperature?

For the plants with aboveground loggers, we have air temperature and soil temperature measured at the same exact location (air temperature measured by the aboveground logger and soil temperature measured with the thermometer). We use these plants to test for correlations between air and soil temperature, using all temperature values.

## Logger data for the whole period:

```{r fig.height=3, fig.width=4, message=FALSE, warning=FALSE}
logger_data%>%
  filter(above_below=="A")%>%
  group_by(logger_nr) %>%
  summarize(mean_logger=mean(temp),temp_term=mean(temp_term))%>%
  ggplot(.,aes(x=temp_term,y=mean_logger))+
  geom_point(size=0.5)+
  geom_smooth(method="lm",size=0.5)+
  geom_abline(lty="dashed",alpha=0.8)+
  xlab("Soil temperature (10 cm depth) measured in May")+
  ylab("Air temperature = Mean temperature\nfrom aboveground logger data\n(whole period)")
logger_data%>%
  filter(above_below=="A")%>%
  group_by(logger_nr) %>%
  summarize(mean_logger=mean(temp),temp_term=mean(temp_term))%>%
  lm(mean_logger~temp_term, data = .) %>%
  summ(.)
logger_data%>%
  filter(above_below=="A")%>%
  group_by(logger_nr) %>%
  summarize(mean_logger=mean(temp),temp_term=mean(temp_term))%>%
  lm(mean_logger~temp_term, data = .) %>%
  linearHypothesis(., "temp_term = 1")
```

The slope is significantly different from 1: differences between soil and air temperature change with soil temperature, being larger at higher soil temperatures.

## Logger data for May:

```{r message=FALSE, warning=FALSE, fig.height=3, fig.width=4}
logger_data%>%
  filter(above_below=="A")%>%
  mutate(month = month(datetime))%>%
  filter(month==5)%>%
  group_by(logger_nr) %>%
  summarize(meanmay_logger=mean(temp),temp_term=mean(temp_term))%>%
  ggplot(.,aes(x=temp_term,y=meanmay_logger))+
  geom_point(size=0.5)+
  geom_smooth(method="lm",size=0.5)+
  geom_abline(lty="dashed",alpha=0.8)+
  xlab("Soil temperature (10 cm depth) measured in May")+
  ylab("Air temperature May = Mean\ntemperature of May from\naboveground logger data")
logger_data%>%
  filter(above_below=="A")%>%
  mutate(month = month(datetime))%>%
  filter(month==5)%>%
  group_by(logger_nr) %>%
  summarize(meanmay_logger=mean(temp),temp_term=mean(temp_term))%>%
  lm(meanmay_logger~temp_term, data = .) %>%
  summ(.)
logger_data%>%
  filter(above_below=="A")%>%
  mutate(month = month(datetime))%>%
  filter(month==5)%>%
  group_by(logger_nr) %>%
  summarize(meanmay_logger=mean(temp),temp_term=mean(temp_term))%>%
  lm(meanmay_logger~temp_term, data = .) %>%
  linearHypothesis(., "temp_term = 1")
```

When looking only at air temperature in May, the slope is also significantly different from 1. But it is not significantly different from zero! So there is no relationship between soil and air temperature.

## Logger data for April-May-June:

```{r message=FALSE, warning=FALSE, fig.height=3, fig.width=4}
logger_data%>%
  filter(above_below=="A")%>%
  mutate(month = month(datetime))%>%
  filter(month==4|month==5|month==6)%>%
  group_by(logger_nr) %>%
  summarize(meanspring_logger=mean(temp),temp_term=mean(temp_term))%>%
  ggplot(.,aes(x=temp_term,y=meanspring_logger))+
  geom_point(size=0.5)+
  geom_smooth(method="lm",size=0.5)+
  geom_abline(lty="dashed",alpha=0.8)+
  xlab("Soil temperature (10 cm depth) measured in May")+
  ylab("Air temperature spring = Mean\ntemperature of April-May-June\nfrom aboveground logger data")
logger_data%>%
  filter(above_below=="A")%>%
  mutate(month = month(datetime))%>%
  filter(month==4|month==5|month==6)%>%
  group_by(logger_nr) %>%
  summarize(meanspring_logger=mean(temp),temp_term=mean(temp_term))%>%
  lm(meanspring_logger~temp_term, data = .) %>%
  summ(.)
logger_data%>%
  filter(above_below=="A")%>%
  mutate(month = month(datetime))%>%
  filter(month==4|month==5|month==6)%>%
  group_by(logger_nr) %>%
  summarize(meanspring_logger=mean(temp),temp_term=mean(temp_term))%>%
  lm(meanspring_logger~temp_term, data = .) %>%
  linearHypothesis(., "temp_term = 1")
```

When looking at air temperature in the period April-May-June, the slope is also significantly different from 1: differences between soil and air temperature change with soil temperature, being larger at higher soil temperatures.

# Q3: Do correlations between soil and air temperature vary with soil temperature?

Or: How useful soil temperature is as a cue for air temperature, i.e., as a cue for spring advancement?

We test this using logger pairs. 

Plot with all paired logger data, one line per logger id

```{r fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
ggplot(logger_data_pairs,aes(x=datetime,y=temp,group=logger_nr))+
  geom_line(size=0.1)+
  geom_vline(xintercept=as.numeric(logger_data_pairs$datetime[2044]), 
             linetype=3,color="red")
```

```{r}
logger_data_pairs%>%
  group_by(logger_nr)%>%
  summarise(pair=first(pair),last_date=last(datetime))%>%
  arrange(pair)
```

Most loggers (131 out of 145) end up on June 6th, so using data until this date.

Plot all logger pairs, to see how each of them looks.

Make one plot for each logger pair and save all as a pdf in output folder.

```{r fig.height=5, fig.width=8, message=FALSE, warning=FALSE, include=FALSE}
plots <- logger_data_pairs %>%
  mutate(month = month(datetime))%>%
         filter(month==4|month==5|month==6)%>%
         filter(datetime<"2018-06-06")%>%
  group_by(pair) %>%
  do(plot = ggplot(data = .) + 
       geom_line(aes(x=datetime,y=temp,color=above_below),size=0.1) +
       ggtitle(paste("pair ",unique(.$pair))))
pdf("output/plots_loggerpairs.pdf",width=12)
plots$plot
dev.off()
```

After looking at the plots, make a new variable pair_problem:
- 0 = no problem
- 1 = some problems (remove)

```{r}
# Define the groups of logger pairs
logger_data_pairs<-logger_data_pairs%>%
  mutate(pair_problem=ifelse(pair%in%c(1:4,11,13:16,21:62,70),0,1))
```

Use only logger pairs with no problems (52 pairs out of 73): pairs 1-4,11,13-16,21-62,70.

We probably should also use only loggers with no problems for the analyses in Q1 and Q2.

## Correlations soil-air temperature over the period April-May-June

For each date and logger pair, calculate mean, max and min of air and soil temperature (from, respectively, the above and belowground logger). Then, calculate the correlation coefficient for air and soil temperatures over the period April-May-June. Finally, regress these correlation coefficients on mean soil temperature (from the belowground logger) **for the same period (April-May-June)**.

Using only data till June 5th, included (as most loggers stop on June 6th).

```{r message=FALSE, warning=FALSE, fig.height=4, fig.width=6}
(logger_data_pairs%>%
     mutate(month = month(datetime))%>%
     filter(month==4|month==5|month==6)%>%
     mutate(date=date(datetime))%>%
     filter(datetime<"2018-06-06")%>%
     filter(!is.na(date))%>%
     filter(pair_problem==0)%>%
     group_by(date,pair,above_below)%>%
     summarise(mean=mean(temp,na.rm=T),max=max(temp),min=min(temp),
               plot=first(plot))%>%
     pivot_wider(names_from="above_below",values_from=c("mean","max","min"))%>%
     group_by(pair)%>%
     summarise(corr_airsoil_mean=cor(mean_A,mean_B,use="pairwise.complete.obs"),
               corr_airsoil_max=cor(max_A,max_B,use="pairwise.complete.obs"),
               corr_airsoil_min=cor(min_A,min_B,use="pairwise.complete.obs"),
               plot=first(plot))%>%
     pivot_longer(cols=corr_airsoil_mean:corr_airsoil_min,
                  names_to="measure",values_to="corr")%>%
     left_join(logger_data_pairs%>%
                 mutate(month = month(datetime))%>%
                 filter(month==4|month==5|month==6)%>%
                 filter(datetime<"2018-06-06")%>%
                 filter(pair_problem==0)%>%
                 filter(above_below=="B")%>%
                 group_by(pair)%>%
                 summarise(meansoiltemp=mean(temp),dist=first(dist))))%>%
  # NAs for plots H11 and H15 --> no Belowground loggers
  ggplot(.,aes(x=meansoiltemp,y=corr,color=measure))+geom_point()+
  geom_smooth(method="lm")+
  xlab("Mean soil temperature for April-May-June")+
  ylab("Correlation coef. air-soil temperature\nover April-May-June\n(based on daily mean/min/max temperatures)")+
  geom_text_repel(data=. %>% filter(corr<0),aes(label=pair))
```

Most of the pairs with negative correlations have disappeared when considering only pairs without problems!

Linear models testing the effect of soil temperature on correlations between soil and air temperature:

```{r message=FALSE, warning=FALSE}
(logger_data_pairs%>%
     mutate(month = month(datetime))%>%
     filter(month==4|month==5|month==6)%>%
     mutate(date=date(datetime))%>%
     filter(datetime<"2018-06-06")%>%
     filter(!is.na(date))%>%
     filter(pair_problem==0)%>%
     group_by(date,pair,above_below)%>%
     summarise(mean=mean(temp,na.rm=T),max=max(temp),min=min(temp),
               plot=first(plot))%>%
     pivot_wider(names_from="above_below",values_from=c("mean","max","min"))%>%
     group_by(pair)%>%
     summarise(corr_airsoil_mean=cor(mean_A,mean_B,use="pairwise.complete.obs"),
               corr_airsoil_max=cor(max_A,max_B,use="pairwise.complete.obs"),
               corr_airsoil_min=cor(min_A,min_B,use="pairwise.complete.obs"),
               plot=first(plot))%>%
     pivot_longer(cols=corr_airsoil_mean:corr_airsoil_min,
                  names_to="measure",values_to="corr")%>%
  left_join(logger_data_pairs%>%
                 mutate(month = month(datetime))%>%
                 filter(month==4|month==5|month==6)%>%
                 filter(datetime<"2018-06-06")%>%
                 filter(pair_problem==0)%>%
                 filter(above_below=="B")%>%
                 group_by(pair)%>%
              summarise(meansoiltemp=mean(temp),dist=first(dist))))%>%
  group_by(measure)%>%
  do(fitcorr = tidy(lm(corr~meansoiltemp, data = .))) %>% 
  unnest(fitcorr)%>%
  kable()
```

All significant.

## Correlations soil-air temperature for May only

For each date and logger pair, calculate mean, max and min of air and soil temperature (from, respectively, the above and belowground logger). Then, calculate the correlation coefficient for air and soil temperatures over the month of May. Finally, regress these correlation coefficients on mean soil temperature (from the belowground logger) **for the same period (May)**.


```{r message=FALSE, warning=FALSE, fig.height=4, fig.width=6}
(logger_data_pairs%>%
  mutate(month = month(datetime))%>%
  filter(month==5)%>%
  mutate(date=date(datetime))%>%
  filter(!is.na(date))%>%
  filter(pair_problem==0)%>%
  group_by(date,pair,above_below)%>%
  summarise(mean=mean(temp,na.rm=T),max=max(temp),min=min(temp),
               plot=first(plot))%>%
  pivot_wider(names_from="above_below",values_from=c("mean","max","min"))%>%
  group_by(pair)%>%
  summarise(corr_airsoil_mean=cor(mean_A,mean_B,use="pairwise.complete.obs"),
            corr_airsoil_max=cor(max_A,max_B,use="pairwise.complete.obs"),
            corr_airsoil_min=cor(min_A,min_B,use="pairwise.complete.obs"),
            plot=first(plot))%>%
   pivot_longer(cols=corr_airsoil_mean:corr_airsoil_min,
                names_to="measure",values_to="corr")%>%
  left_join(logger_data_pairs%>%
              mutate(month = month(datetime))%>%
              filter(month==5)%>%
              filter(pair_problem==0)%>%
              filter(above_below=="B")%>%
              group_by(pair)%>%
              summarise(meansoiltemp=mean(temp),dist=first(dist))))%>%
  # NAs for plots H11 and H15 --> no Belowground loggers
  ggplot(.,aes(x=meansoiltemp,y=corr,color=measure))+geom_point()+
  geom_smooth(method="lm")+
  xlab("Mean soil temperature for May")+
  ylab("Correlation coef. air-soil temperature\nover May\n(based on daily mean/min/max temperatures)")+
  geom_text_repel(data=. %>% filter(corr<0),aes(label=pair))
```

Linear models testing the effect of soil temperature on correlations between soil and air temperature:

```{r message=FALSE, warning=FALSE}
(logger_data_pairs%>%
  mutate(month = month(datetime))%>%
  filter(month==5)%>%
  mutate(date=date(datetime))%>%
  filter(!is.na(date))%>%
  filter(pair_problem==0)%>%
  group_by(date,pair,above_below)%>%
  summarise(mean=mean(temp,na.rm=T),max=max(temp),min=min(temp))%>%
  pivot_wider(names_from="above_below",values_from=c("mean","max","min"))%>%
  group_by(pair)%>%
  summarise(corr_airsoil_mean=cor(mean_A,mean_B,use="pairwise.complete.obs"),
            corr_airsoil_max=cor(max_A,max_B,use="pairwise.complete.obs"),
            corr_airsoil_min=cor(min_A,min_B,use="pairwise.complete.obs"))%>%
   pivot_longer(cols=corr_airsoil_mean:corr_airsoil_min,
                names_to="measure",values_to="corr")%>%
  left_join(logger_data_pairs%>%
              mutate(month = month(datetime))%>%
  filter(month==5)%>%
  filter(pair_problem==0)%>%
  filter(above_below=="B")%>%
  group_by(pair)%>%
  summarise(meansoiltemp=mean(temp),dist=first(dist))))%>%
  group_by(measure)%>%
  do(fitcorr = tidy(lm(corr~meansoiltemp, data = .))) %>% 
  unnest(fitcorr)%>%
  kable()
```

All significant.

# Session Info

```{r include=FALSE}
sessionInfo()
```
